\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-05-22 doc/paper.Rmd
proofing up to coarseness

Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. 
\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail using CS1. The highest percentage of failed tests under CS2 is 1.3% for different-source tests using a validation window size $w_v$ of 60 pixels.  \hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }\hh{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }

@@

```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2). Observed failure rates are higher than expected rates."}
allinv$method <- "CS2"
allp$method <- "CS1"

@@

prof_fails <- rbind(
  allinv %>% dplyr::select(wo, wv, method, p_value, match),
  allp %>% filter(coarse == 0.25) %>% 
    dplyr::select(wo, wv, method, p_value, match)
)
fails <- prof_fails %>%
  group_by(wo, wv, method, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
)

fails %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(aes(size="Observed")) +
  xlab("Size of validation window in pixels") +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13)) +
  facet_wrap(~method, labeller="label_both") +
  geom_point(aes(y = failed*100, size = "Expected"), data=simusumm) +
  scale_size_manual("Failures", values=c(1.5,3))
```

>>>>

As described above, the Chumbley-score is based on three parameters: coarseness $c$ and the sizes of the optimization window $w_o$ and validation window $w_v$. 
In a first run of results, we applied default settings for the parameters, as suggested in  \citet{hadler}: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$, and varied the size of the  validation window $w_v$ in steps of 10  from 10 pixels to 60 pixels.
Based on a significance level $\alpha$ of 0.05 for the test, this results in a correct identification of same-source and different-source toolmarks of 93.5\% to 94.1\%, corresponding to a rate of false negatives between 0.28 and 0.36 and a rate of false positives between 0.05 and 0.06. 
However, the most prominent result we encountered, are the high number of failed tests, i.e. the number of instances, in which CS1 did not return any result.
\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail using CS1. The highest percentage of failed tests under CS2 is 1.3% for different-source tests using a validation window size $w_v$ of 60 pixels.  Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.25 resulting in a distribution of optimal locations different from the assumed uniform.

@@

```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2). Observed failure rates of (CS1) are higher than expected rates."}
allinv$method <- "CS2"
allp$method <- "CS1"

@@

prof_fails <- rbind(
  allinv %>% dplyr::select(wo, wv, method, p_value, match),
  allp %>% filter(coarse == 0.25) %>% 
    dplyr::select(wo, wv, method, p_value, match)
)
stats <- prof_fails %>%
  group_by(wo, wv, method) %>% summarize(
  failed = sum(is.na(p_value)),
  correctplus = sum(p_value<=0.05 & match, na.rm=TRUE),
  correctminus = sum(p_value > 0.05 & !match, na.rm=TRUE),
  n=n(),
  correct = (correctplus+correctminus)/(n-failed)
)
#stats %>% 
#  ggplot(aes(x = wv, y = (correctplus+correctminus)/(n-failed))) +
#  geom_point(aes(colour=method))
#range(subset(stats, method=="CS1")$correct)
#[1] 0.9347536 0.9410993

fails <- prof_fails %>%
  group_by(wo, wv, method, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n(),
  falseneg = sum(p_value > 0.05 & match, na.rm=TRUE),
  falsepos = sum(p_value <= 0.05 & !match, na.rm=TRUE)
)
fnr <- with(subset(fails, method=="CS1"), falseneg/(n-failed))
#summary(fnr[fnr>0])
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.2819  0.2882  0.2965  0.3056  0.3075  0.3623 
fpr <- with(subset(fails, method=="CS1"), falsepos/(n-failed))
#summary(fpr[fpr>0])
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#0.05468 0.05588 0.05855 0.05852 0.06120 0.06227

fails %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(aes(size="Observed")) +
  xlab(expression("Size of validation window "~w[v]~" in pixels")) +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13)) +
  facet_wrap(~method, labeller="label_both") +
  geom_point(aes(y = failed*100, size = "Expected"), data=simusumm) +
  scale_size_manual("Failures", values=c(1.5,3))
```

### 2018-05-22 doc/paper.tex
proofing up to coarseness

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
\end{figure}

@@

Both algorithms (CS1) and (CS2) are implemented in the R package
\texttt{toolmaRk} \citep{toolmark}. We applied both methods to all
pairwise land-to-land comparisons of the Hamby scans provided by NIST
for a total of 85,491 land-to-land comparisons.

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1. The highest percentage of failed tests under CS2
is 1.3\% for different-source tests using a validation window size
\(w_v\) of 60 pixels.
{\textcolor{orange}{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }}{\textcolor{orange}{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }}

>>>>

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
\end{figure}

@@

Both algorithms (CS1) and (CS2) are implemented in R \citep{R}. (CS1) is
available from package \texttt{toolmaRk} \citep{toolmark}, (CS2) is
available from a modified version of the \texttt{toolmaRk} package
available from GitHub (\url{https://github.com/heike/toolmaRk}). We
applied both methods to all pairwise land-to-land comparisons of the
Hamby scans provided by NIST for a total of 85,491 land-to-land
comparisons.

@@

As described above, the Chumbley-score is based on three parameters:
coarseness \(c\) and the sizes of the optimization window \(w_o\) and
validation window \(w_v\). In a first run of results, we applied default
settings for the parameters, as suggested in \citet{hadler}:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\), and varied the size of
the validation window \(w_v\) in steps of 10 from 10 pixels to 60
pixels. Based on a significance level \(\alpha\) of 0.05 for the test,
this results in a correct identification of same-source and
different-source toolmarks of 93.5\% to 94.1\%, corresponding to a rate
of false negatives between 0.28 and 0.36 and a rate of false positives
between 0.05 and 0.06. However, the most prominent result we
encountered, are the high number of failed tests, i.e.~the number of
instances, in which CS1 did not return any result. \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1. The highest percentage of failed tests under CS2
is 1.3\% for different-source tests using a validation window size
\(w_v\) of 60 pixels. Rates of expected failures are based on simulation
runs using covariances between locations of same-source profiles of
0.854, and 0.120 for locations from different-source profiles, matching
observed covariances for the Hamby scans. Observed failure rates are
higher than expected. This might be due to remaining sub-class structure
at a coarseness of 0.25 resulting in a distribution of optimal locations
different from the assumed uniform.
